---
title: "Advanced Data Analysis | Final Group Project | Spring 2018"
output: html_notebook
date: "April 25, 2018"
author: "Jasmine Chiaka Chukwueke" <JChukwueke@wustl.edu>"; Keenya Frazier <frazierks@wustl.edu>; Chaitanay Siddagunta <c.siddag@wustl.edu>
---

'# PROLOG   ################################################################'

'# PROJECT: Advanced Data Analysis | Final Group Project
'# PURPOSE: FINAL PROJECT RCODE for COURSE SPECIAL TOPICS: ADVANCED DATA ANALYSIS
            Poisson and Negative Binomial regression. and use of R markdown

'# DIR:     https://wustl.blackboard.com/webapps/blackboard/execute/content/file?cmd=view|content_id=_3357249_1|course_id=_189741_1|framesetWrapped=true

'# DATA:    Adult.2.csv | peds.csv

'# AUTHOR:  Chukwueke, Jasmine C; Fraizer Keenya; Chaitu Siddagunta
'# CREATED: April 18, 2018
'# LATEST:  April 23, 2018
'# NOTES:   # Final Group Project | RCode
            # Reference for ggplot2: http://www.cookbook-r.com/Graphs/
            # Reference for Margins pkg: http://cran.us.r-project.org
            # Reference for Stargazer pkg: http://cran.us.r-project.org
'# PROLOG   ###############################################################'


#### Load the libraries

```{r}
# install.packages("margins",repos = "http://cran.us.r-project.org") 
# install.packages("stargazer",repos = "http://cran.us.r-project.org")
library(haven) #read dta file 
library(MASS) #Negative binomial regression 
library(lmtest) #model comparison 
library(stargazer) #models presentation 
library(sandwich)  #robust  
library(margins) #Marginal effects 
library(ggplot2) #Graphs 
library(readxl) # excel file
library(tidyverse)
```

#### Read data(Adult Patients)

```{r}
adult <- read_csv("~/Dropbox/WashU | MPH/Learning | In St. Louis/2017 - 2018/Spring 2018/Advanced Data Analysis/Final Group Project/Adult.2.csv")
adult$Age <- "Adult"
View(adult)
```

#### Read data(Pediatric Patients)

```{r}
peds <- read.csv("~/Dropbox/WashU | MPH/Learning | In St. Louis/2017 - 2018/Spring 2018/Advanced Data Analysis/Final Group Project/peds.csv")
View(peds)
peds$SIR <- NA
peds$Age <- "Child"
```

#### Merge two datasets

```{r}
Patients <- rbind(peds, adult)
View(Patients)
```

#### Data management on the Dataset

```{r}
str(Patients)
Patients
#Code to Crete two new "Age" Variables: One character and One factor:
Patients$Age.f <- as.factor(Patients$Age)
str(Patients)
table(Patients$Age.f)
Patients$Age_cat[Patients$Age=="Child"] <- 0
Patients$Age_cat[Patients$Age=="Adult"] <- 1
str(Patients$Age_cat)
Patients$Age_cat <- as.character(Patients$Age_cat)
Patients$Age.f <- as.factor(Patients$Age_cat)
table(Patients$Age.f)
class(Patients$Age.f)
table(Patients$Age)
class(Patients$Age)

#Create a New Variable: Infection_cat
table(Patients$Infection_Count)
Patients$Infection_cat <- NA
str(Patients)
Patients$Infection_cat[Patients$Infection_Count==0] <- 0
Patients$Infection_cat[Patients$Infection_Count> 0] <- 1
str(Patients)
View(Patients)
#Create a New Variable: Infection.f
Patients$Infection.f <- as.factor(Patients$Infection_cat)
str(Patients)
# 7953 Rows | 17 Columns

#Change event to numeric:
Patients$Infection.f<- as.numeric(Patients$Infection.f)

#View Operative Procedures
table(Patients$Operative_Procedure)

Patients$OperativeProcedure[Patients$Operative_Procedure==0] <- 0
Patients$Infection_cat[Patients$Infection_Count> 0] <- 1

```

#### Recode Opeartive_Procedure variable as a factor variable and label it

```{r}

##Recode Opeartive_Procedure variable as a factor variable and label it
Patients$Surgery.f[
  Patients$Operative_Procedure=="Appendix surgery"| Patients$Operative_Procedure=="Bile duct, liver or pancreatic surgery" | Patients$Operative_Procedure=="Colon surgery " | Patients$Operative_Procedure=="Exploratory abdomial surgery (laparotomy)" | Patients$Operative_Procedure=="Gallbladder surgery" | Patients$Operative_Procedure=="Hysterectomy,abdominal" | Patients$Operative_Procedure=="Rectal surgery" | Patients$Operative_Procedure=="Small bowel surgery" | Patients$Operative_Procedure=="Exploratory abdominal surgery (laparotomy)" | Patients$Operative_Procedure=="Gastric surgery" | Patients$Operative_Procedure=="Hysterectomy, abdominal" | Patients$Operative_Procedure=="Spleen surgery" | Patients$Operative_Procedure=="Kidney surgery"]<-0

Patients$Surgery.f[
  Patients$Operative_Procedure=="Cardiac surgery"| Patients$Operative_Procedure=="Thoracic surgery" | Patients$Operative_Procedure=="Abdominal aortic aneurysm repair" | Patients$Operative_Procedure=="Coronary bypass, chest and donor incisions" | Patients$Operative_Procedure=="Coronary bypass, chest incision only" | Patients$Operative_Procedure=="Pacemaker surgery"] <- 1

Patients$Surgery.f[
  Patients$Operative_Procedure=="Cesarean section" | Patients$Operative_Procedure=="Hysterectomy, vaginal" | Patients$Operative_Procedure=="Ovarian surgery"] <- 2

Patients$Surgery.f[
  Patients$Operative_Procedure=="Kidney transplant"| Patients$Operative_Procedure=="Liver transplant " | Patients$Operative_Procedure=="Heart transplant "] <- 3

Patients$Surgery.f[
  Patients$Operative_Procedure=="Hip prosthesis"| Patients$Operative_Procedure=="Knee prosthesis" | Patients$Operative_Procedure=="Laminectomy" |Patients$Operative_Procedure=="Open reduction of fracture"| Patients$Operative_Procedure=="Spinal fusion"] <- 4

table(Patients$Surgery.f)
class(Patients$Surgery.f)

Patients$Surgery.f <- factor(Patients$Surgery.f,
  levels = c(0,1,2,3,4),
  labels = c("Abdominal", "Thoracic", "Pelvic", "Transplant", "MSK"))
class(Patients$Surgery.f)
table(Patients$Surgery.f)

#What should be the reference?
#Pelvic and MSK = Same
#Abdominal, Thoracic, Transplant = Different
```

#### Run a Poisson regression model for age category, type of surgery performed,and SSIs as the dependent variable (remember the predictors as factor variables).

```{r}
str(Patients)
modP<- glm(as.numeric(Infection.f) ~ factor(Age.f)+factor(Surgery.f) + Procedure_Count, family="poisson", data=Patients)
summary(modP)
```

#### Run a Negative Binomial regression model for age, type of surgery performed as predictors and SSIs as the dependent variable.

```{r}
modN<- glm.nb(Infection.f ~ factor(Age.f) + factor(Surgery.f) + Procedure_Count, data=Patients)
summary(modN)
```

#### Compare models using the likelihood ratio test. Which model is better and on what basis do you make that conclusion?

```{r}
lrtest(modP, modN)
```

```{r}
stargazer(modP, modN, title="Model Comparison", type="text",align=TRUE,single.row=TRUE)
```

#### Compute robust standard errors and calculate incidence rate ratios and 95% CIs using the robust standard errors

```{r}
robust<-coeftest(modN, vcov = sandwich) 
robust
```

####Incidence rate ratios and 95%CIs

```{r}
est <- cbind(IRR = coef(modN), "2.5%"=robust[,1]-1.96*robust[,2], "97.5%"=robust[,1]+1.96*robust[,2])
exp(est)
```

#### How well did our models do at predicting 0's

```{r}
mu <- predict(modP, type = "response") # predict expected mean count
exp <- sum(dpois(x = 0, lambda = mu)) # sum the probabilities of a 0 count for each mean
round(exp) # predicted number of 0's
sum(Patients$Infection.f < 1) # observed number of 0's
```

#### number of counts of SSIs in the last 30 days

```{r}
p <- ggplot(data=Patients)
p <- p + geom_bar(aes(x=Surgery.f))
p <- p + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p

```

[End of Code]
